Tree establishment in models of forest dynamics: a key priority for further research

### 2022-12-15 manuscript/manuscriptRW.Rmd
couple of changes in the discussion section

---
title: "Tree establishment in models of forest dynamics: a key priority for further research"
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Tim Anders,
- Kristin H. Braziunas,
- Josef	Bruna,
- Samuel Fischer,
- Jessica Hetzer,
- Thomas Hickler,
- Christian	Hochauer,
- Manfred Lexer,
- Heike	Lischke,
- Mats Mahnken,
- Paola	Mairota,
- Katarina Merganicova,
- Tobias Mette,
- Xavier Morin,
- Werner Rammer,
- Simon	Scheiter,
- Daniel Scherrer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: global-change-biology.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Fifteen models of forest dynamics (including two model featuring
variants) were used to assess tree establishment. They simulate forest
dynamics at the stand, landscape and global scale (Table
\@ref(tab:tableModels)). The approaches used for model construction and
their origin differ strongly, with most of the models featuring a
largely "process-based" approach, whereas two models are based on
formulations derived from the statistical analysis of inventory data
(SYBILA and xComp). Some of the models largely rely on the original
approach underlying forest gap models (e.g., ForClim 1 and
LandClim)[,]{.underline} whereas some [a]{.underline}re based on plant
ecophysiological processes (e.g., FORMIND and iLand). The different
approaches underlying the models have strongly influenced the
formulation of tree establishment.

The establishment modules can be categorized as 'regeneration' or
'recruitment' models [@vanclay1994; @konig_tree_2022] (see Table
\@ref(tab:tableModels)). Regeneration models include processes such as
flowering and pollination, seed production, seed dispersal, germination
seedling [g]{.underline}rowth, which ultimately lead to the simulated
number of trees. Recruitment models, in contrast, introduce a certain
number of new trees with certain characteristics such as biomass or
diameter, without explicitly considering earlier development processes.
We can further distinguish models that feature a feedback in the
establishment, i.e., where the level and species composition of
establishment is influenced by the existing adult tree community via the
production of seeds, seedlings or saplings, from models that do not
contain such feedback.

The starting point for in the models differs as well, ranging from
seedbank, seed or seedling to sapling (i.e., trees much larger than 10
cm height, often at least 2-3 m tall).This implies that the larger the
size threshold, the larger the number of ecological processes that must
be aggregated via parameterizations because they cannot be treated
explicitly. This reduces model complexity but comes at the cost of
blurring process representation.

The overall complexity in the establishment modules also varies
considerably. Following @bugmann2022, we can classify seven models
(iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as
having rather high complexity in their establishment modules (average
complexity across all processes \>0, Table \@ref(tab:tableModels)). In
most models[,]{.underline} the two processes in the establishment
modules with the highest complexity are the statistical approach used to
define the successful number of new treesand how frost occurrence is
taken into account.

Two models, ForClim and ForCEEPs, were used here with two alternative
variants. ForClim variant 1 [@bugmann1996] is based on a recruitment
module that adheres closely to the concept introduced by @botkin1972,
whereas ForClim variant 11 is adopting a rather different approach that
considers space filling along with different 'points of entry' for the
action of specific environmental factors [@huber2020]. These two model
variants allow us to evaluate a more process-based module (variant 11)
against a simple module (variant 1), while the rest of the model
structure is identical. Similarly, the two ForCEEPS variants allow us to
isolate the importance of the canopy feedback via seed trees for the
quantity and quality of simulated ingrowth[,]{.underline} as one variant
includes this feedback whereas the other does not.

@@

kable(modelDF, 
      #format = "markdown",
      format = "latex", booktabs = TRUE,
      caption = "List of models included.") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |> 
      kableExtra::column_spec(column = 2, width = "3em") |> 
      kableExtra::column_spec(column = 3, width = "3em") |> 
      kableExtra::column_spec(column = 4, width = "4em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "5em") |>
      kableExtra::column_spec(column = 7, width = "5em") |>
      kableExtra::column_spec(column = 8, width = "8em") |>
      kableExtra::column_spec(column = 9, width = "8em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "5em") |>
      kableExtra::column_spec(column = 12, width = "5em") |>
      kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
                                 "PB-Process based, EM-Empirical", 
                                 "C-Cohort, I-Individual",
                                 "RG-Regeneration, RC-Recruitment",
                                 "Y-Yes, N-No",
                                 "D-Daily, M-Monthly"),

@@

Ingrowth data covering a wide range of environmental conditions are hard
to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number sites, to evaluate how well ingrowth is
captured. The observations used here are from a novel and unprecedented
network of sites in forest reserves that represent the range of
environmental gradients in temperature and precipitation in central
Europe as compiled in the framework of the EuFoRIa network[@euforia2019]
(Figure \@ref(fig:map)). All these forests have not been managed for at
least 10 years prior to their designation as reserves, and most of them
provide time series of natural forest dynamics over multiple decades and
up to ca. 80 years. The census periods range from 3 to 37 years, with an
average of 14 years. The data provide information at the tree level,
thus allowing for sequential comparison such as individual-tree ingrowth
and death between the measurements.

@@

The overarching goal of the experiments was to assess the regeneration
niche as it arises from empirical data against its representation in a
wide range of models of forest dynamics. We define the regeneration
niche as the passing of a diameter threshold of 7 or 10
cm[,]{.underline} respectively (synonym: ingrowth). To this end, each
modeling group was provided with a detailed protocol (cf. Supplementary
Material 1, henceforth "SM") with instructions how to perform the
simulations, providing input variables on climate and soil conditions,
and the expected output variables. Neither were further site information
(except for the data specified below) nor any data on tree ingrowth
provided prior to the simulation. That is, the models were run in "blind
flight" mode.

The input variables were collected and aggregated to be adapted to the
needs of the different models. Time series of climatic variables from
1981 to 2018 were provided in hourly [Era5-land data, Muñoz Sabater
-@muñozsabater2019], daily and monthly resolution [CHELSA data, Karger
et al. -@karger2021]. Some variables required for some models such as
relative humidity or vapor pressure deficit had to be calculated from
these variables. The final strategy how to use the climate data was
different in each model, based on the approach that best suited the
model (Table \@ref(tab:tableModels)). Soil quality data was provided as
a continuous value between 1 and 5 [Soilgrids dataset, Hengl et al.
-@hengl2017, Poggio et al. -@poggio2021]. The protocol also provided the
elevation, slope and aspect for each of the 200 sites
[@asterscienceteam2019], but no other spatial information such as
coordinates was provided, with the exception of iLand[,]{.underline}
which used the blurred coordinates to derive highly detailed soil data.

The simulations were run in the absence of natural disturbances. Only
the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE [ONE
SENTENCE]{.underline} OF WHY YOU INCLUDED DISTURBANCES AND WHAT TYPE.
The simulations were set up to sample species-specific ingrowth rates
per decade and per ha in the equilibrium state of the model, typically
entailing a "spin-up" run (as we did not provide any forest data). The
modeling teams decided on the simulated area and how they derived these
samples. The exact length of the simulation was also decided by the
modeling teams (Table \@ref(tab:tableModels)). Further details on how
each modeling team prepared the simulations and the outputs are
available in SM 2.

@@

The simulations were run for mixed species forests (not multiple
single-species simulations) using mixtures of eleven species or genera
for which ingrowth data of sufficient quality were available from
EuFoRIa: *Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus,
Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior,
Quercus spp., Alnus glutinosa, Pinus sylvestris.* Two models included
their standard set of species for the simulations which is much larger
(i.e., ForClim 1, ForClim 11). This also applies to TreeMig, but in this
model *Betula pendula* was not included because HERE **TREEMIG**
PROVIDES THE END OF THIS SENTENCE SAYING WHY B pendula WAS NOT INCLUDED.
In three models, fewer than these elveven species were simulated (4C,
xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). HERE **4C** PROVIDES
[ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS SPECIES. HERE
**XCOMP** PROVIDES [ONE SENTENCE]{.underline} WHY THEY SIMULATED LESS
SPECIES. HERE **LPJ-GUESS** PROVIDES [ONE SENTENCE]{.underline} WHY THEY
SIMULATED LESS SPECIES. Finally, the model aDGVM2 is based on a suite of
broad Plant Functional Types rather than individual species, and hence
this model used its default approach of PFTs for the simulations.

@@

Three models did not provide results from all the simulated samples or
sites to avoid unrealistic results, as follows. In the model 4C, a
threshold of a maximum basal area of 100 $m^2 ha^{-1}$ was used to avoid
unrealistic stand basal area data, and therefore not all the samples and
sites were reported. The reason for this is that 4C is not suitable for
long-term simulations without management PLEASE **4C** VERIFY THIS IS
WHAT YOU DID. LPJ-GUESS had 2% of the sample outputs as grasslands that
were considered in the data as zero stand basal area and zero ingrowth
for all the tree species simulated. LPJ-GUESS could not simulate one of
the sites HERE **LPJ-GUESS** PROVIDES [ONE SENTENCE]{.underline} WHY
THEY DID NOT PROVIDE ALL THE SITES. Landis-II missed 0.2% of the
observations HERE **LANDIS-II** PROVIDES [ONE SENTENCE]{.underline} WHY
THEY DID NOT PROVIDE ALL THE SAMPLES/observations.

@@

Both observations and simulated data had no ingrowth in some samples and
at some sites (Table S1). The observed data had slightly more than 3% of
the samples with no ingrowth. Only three models had a larger proportion
of no ingrowth (4C, Landis-II and aDVM2). Three models always simulated
ingrowth for both the 7 and 10 cm threshold (TreeMig and LPJ-GUESS),
i.e., they did not feature any zero values. The other models had a very
low percentage of samples with no ingrowth (0.01% - 1.31%), i.e., they
had distinctly fewer occurrences of zeros compared to the observations.
There was no significant relationship between the mean complexity of the
establishment module (Table \@ref(tab:tableModels)) and the
overestimation proportion at a diameter threshold of 7 cm (Table S2).

@@

For both the observed and the simulated data, and both ingrowth
thresholds, species diversity was lower in the ingrowth compared to
stand-level diversity (Figures S3 and S4). Yet, models simulating higher
diversity at the stand level featured higher diversity in ingrowth
(Figure S3 and S4). Most models captured reasonably well or
overpredicted species diversity at the stand level (Figure
\@ref(fig:combinedRichness)A, center; cf. Figure S5)[,]{.underline} but
a characteristic feature was that simulations had a much lower
variability of diversity compared to observations. Models overpredicting
species richness included ForClim1, ForClim 11, FORMIND and TreeMig
(Figure \@ref(fig:combinedRichness)A, left). Several models
underpredicted species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C
(Figure \@ref(fig:combinedRichness)A, right). The majority of the models
captured reasonably well the ingrowth diversity at lower values of
diversity, but most of them strongly underestimated ingrowth diversity
for higher observed diversity levels (Figure
\@ref(fig:combinedRichness)B, cf. Figure S6). Both the models with
feedback and those without feedback from the adult trees to ingrowth
(via seed production) underestimated ingrowth species diversity compared
to the observations. Again, simulated diversity was considerably less
variable than observed diversity (Fig. \@ref(fig:combinedRichness)B).

@@

Second, there was some congruence in the simulation results by species
across the models. The share in the ingrowth by Silver fir (*Abies
alba*) was captured well across the gradient by most models except for
xComp, PICUS, LandClim and Landis-II, which overestimated the ingrowth
share of this species. There were also some peculiarities evident for
some models, such as 4C that did not simulate *Abies alba*. *Fagus
sylvatica* ingrowth was underestimated by many models across the
environmental gradients, including ForCEEPS(f) and FORMIND, which
featured negligible ingrowth at most sites for this species. Most of the
models overestimated the share of *Picea abies* ingrowth in the cold-wet
part of the gradients. *Pinus sylvestris* and *Quercus* spp. represented
a smaller share of the ingrowth in the observed data mostly at warm-dry
sites. Many models, however, erroneously featured ingrowth for these two
species along most of the climatic gradients although at a low basal
area share. Five models (4C, FORMIND, SYBILA, xComp, and LandClim) had
almost no ingrowth of *Pinus sylvestris* and *Quercus* spp., or did not
feature any ingrowth of of these species at all.

@@

The species diversity of the entire stand was better captured by the
models than the diversity of tree ingrowth, and only a few models
overpredicted stand level diversity. Ingrowth composition defined as
being proportional to adult abundance, regardless of productivity, might
be a simple and cautious assumption to safely maintain relative species
abundances [@hanbury-brown2022]. However, based on the simulation
results, we did not see that models with feedback captured better the
ingrowth species diversity than those without feedback. The similar
performance between models with and without feedback might be because
the models are giving more importance to define the regeneration niche
via abiotic and biotic filters than with the habitat niche of the adult
trees [@grubb1977].

@@

The EuFoRIa data as used in this study is unique and highly suited to
better understand tree establishment. However, three aspects of this
data may also represent considerable limitations. First, we made a
comparison of tree ingrowth in a posited equilibrium state, but we
cannot assess how close the empirical data were to such a state. The
data were collected in forest reserves where no management has taken
place for long periods of time. This makes our assumption of an
equilibrium between forest properties and environmental drivers less
haphazard than it might appear at first sight. However, it may explain
why simulated ingrowth rates were higher than the observed ones. Second,
the data were collected for ingrowth above a 7 cm threshold, thus
limiting the assessment of tree establishment in a holistic sense. This
constituted a hard limit based on which we can understand some aspects
of tree establishment, which in its entirety often comprises a rather
long period since germination [@price2001]. However, many environmental
constraints are acting on young trees [@käber2021] that we were unable
to assess. Yet, ingrowth data with lower thresholds are simply not
available in a harmonized manner across large environmental gradients.
Third, the empirical data were collected from rather small plots that
ranged from 0.2 to up to 1 ha, while we sampled simulated ingrowth rates
from 1 ha areas, which may lead to an incorrect representation of space.
Even though this strategy it is not ideal, it represents a common
challenge when harmonizing diverse data sources originating from varied
sampling strategies [@portier2022]. It would have been extremely
challenging for such a variety of models to follow a protocol where the
spatial sampling size was different at each of the 200 sites.

The design of our sampling protocol did not include spatial aspects. We
considered a wide range of models of forest dynamics, from stand to
global scale. However, the simulation setup was limited to 1 ha ingrowth
values. The lack of spatial considerations potentially puts landscape
models at a disadvantage, as they have been built to be accurate~~l~~ at
the landscape level. Without spatial context, we are limiting ingrowth
to the seed influx from the stand itself, if the model does not have a
background seed input. However, Global Dynamic Vegetation Models should
not be in disadvantage due to the limited spatial considerations [but
cf. Snell et al. -@snell2014] as they are usually lacking dispersal
between cells as they are based on an abstraction of horizontal space,
or even the trees themselves [@hanburybrown_forest_2022].

@@

Models of forest dynamics are important tools in science and decision
support, and how tree establishment is defined has strong implications
for simulated forest properties. The 15 models and variants used here
are facing similar challenges when representing tree establishment. They
generally overestimate tree establishment rates, and the simulated
establishment niche is not always captured accurately as a function of
the biotic (light) and abiotic (temperature and moisture). However, most
models properly capture the diversity of the initial tree populations,
although changes in their formulations including feedback from adult
trees did not have an effect on better capturing the species composition
of ingrowth.

>>>>

---
title: "Tree establishment in models of forest dynamics: a key priority for further research"
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Tim Anders,
- Kristin H. Braziunas,
- Josef	Bruna,
- Samuel Fischer,
- Jessica Hetzer,
- Thomas Hickler,
- Christian	Hochauer,
- Manfred Lexer,
- Heike	Lischke,
- Mats Mahnken,
- Paola	Mairota,
- Katarina Merganicova,
- Tobias Mette,
- Xavier Morin,
- Werner Rammer,
- Simon	Scheiter,
- Daniel Scherrer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
nocite: |
  @reyer, @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @seidl2012a, @lischke2006a, @schumacher2006, @scheiter2009, @scheller2004, @morin2021
linkcolor: blue
csl: global-change-biology.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Fifteen models of forest dynamics (including two model featuring
variants) were used to assess tree establishment. They simulate forest
dynamics at the stand, landscape and global scale (Table
\@ref(tab:tableModels)). The approaches used for model construction and
their origin differ strongly, with most of the models featuring a
largely "process-based" approach, whereas two models are based on
formulations derived from the statistical analysis of inventory data
(SYBILA and xComp). Some of the models largely rely on the original
approach underlying forest gap models (e.g., ForClim 1 and LandClim),
whereas some are based on plant ecophysiological processes (e.g.,
FORMIND and iLand). The different approaches underlying the models have
strongly influenced the formulation of tree establishment.

The establishment modules can be categorized as 'regeneration' or
'recruitment' models [@vanclay1994; @konig_tree_2022] (see Table
\@ref(tab:tableModels)). Regeneration models include processes such as
flowering and pollination, seed production, seed dispersal, germination
seedling growth, which ultimately lead to the simulated number of trees.
Recruitment models, in contrast, introduce a certain number of new trees
with certain characteristics such as biomass or diameter, without
explicitly considering earlier development processes. We can further
distinguish models that feature a feedback in the establishment, i.e.,
where the level and species composition of establishment is influenced
by the existing adult tree community via the production of seeds,
seedlings or saplings, from models that do not contain such feedback.

The starting point for in the models differs as well, ranging from
seedbank, seed or seedling to sapling (i.e., trees much larger than 10
cm height, often at least 2-3 m tall). This implies that the larger the
size threshold, the larger the number of ecological processes that must
be aggregated via parameterizations because they cannot be treated
explicitly. This reduces model complexity but comes at the cost of
blurring process representation.

The overall complexity in the establishment modules also varies
considerably. Following @bugmann2022, we can classify seven models
(iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and TreeMig) as
having rather high complexity in their establishment modules (average
complexity across all processes \>0, Table \@ref(tab:tableModels)). In
most models, the two processes in the establishment modules with the
highest complexity are the statistical approach used to define the
successful number of new trees and how frost occurrence is taken into
account.

Two models, ForClim and ForCEEPs, were used here with two alternative
variants. ForClim variant 1 [@bugmann1996] is based on a recruitment
module that adheres closely to the concept introduced by @botkin1972,
whereas ForClim variant 11 is adopting a rather different approach that
considers space filling along with different 'points of entry' for the
action of specific environmental factors [@huber2020]. These two model
variants allow us to evaluate a more process-based module (variant 11)
against a simple module (variant 1), while the rest of the model
structure is identical. Similarly, the two ForCEEPS variants allow us to
isolate the importance of the canopy feedback via seed trees for the
quantity and quality of simulated ingrowth, as one variant includes this
feedback whereas the other does not.

@@

kable(modelDF, 
      #format = "markdown",
      format = "latex", booktabs = TRUE,
      caption = "List of models included.") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |> 
      kableExtra::column_spec(column = 2, width = "3em") |> 
      kableExtra::column_spec(column = 3, width = "3em") |> 
      kableExtra::column_spec(column = 4, width = "4em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "5em") |>
      kableExtra::column_spec(column = 7, width = "5em") |>
      kableExtra::column_spec(column = 8, width = "8em") |>
      kableExtra::column_spec(column = 9, width = "8em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "5em") |>
      kableExtra::column_spec(column = 12, width = "8em") |>
      kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
                                 "PB-Process based, EM-Empirical", 
                                 "C-Cohort, I-Individual",
                                 "RG-Regeneration, RC-Recruitment",
                                 "Y-Yes, N-No",
                                 "D-Daily, M-Monthly"),

@@

Ingrowth data covering a wide range of environmental conditions are hard
to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number sites, to evaluate how well ingrowth is
captured. The observations used here are from a novel and unprecedented
network of sites in forest reserves that represent the range of
environmental gradients in temperature and precipitation in central
Europe as compiled in the framework of the EuFoRIa network
[@euforia2019] (Figure \@ref(fig:map)). All these forests have not been
managed for at least 10 years prior to their designation as reserves,
and most of them provide time series of natural forest dynamics over
multiple decades and up to ca. 80 years. The census periods range from 3
to 37 years, with an average of 14 years. The data provide information
at the tree level, thus allowing for sequential comparison such as
individual-tree ingrowth and death between the measurements.

@@

The overarching goal of the experiments was to assess the regeneration
niche as it arises from empirical data against its representation in a
wide range of models of forest dynamics. We define the regeneration
niche as the passing of a diameter threshold of 7 or 10 cm, respectively
(synonym: ingrowth). To this end, each modeling group was provided with
a detailed protocol (cf. Supplementary Material 1, henceforth "SM") with
instructions how to perform the simulations, providing input variables
on climate and soil conditions, and the expected output variables.
Neither were further site information (except for the data specified
below) nor any data on tree ingrowth provided prior to the simulation.
That is, the models were run in "blind flight" mode.

The input variables were collected and aggregated to be adapted to the
needs of the different models. Time series of climatic variables from
1981 to 2018 were provided in hourly [Era5-land data, Muñoz Sabater
-@muñozsabater2019], daily and monthly resolution [CHELSA data, Karger
et al. -@karger2021]. Some variables required for some models such as
relative humidity or vapor pressure deficit had to be calculated from
these variables. The final strategy how to use the climate data was
different in each model, based on the approach that best suited the
model (Table \@ref(tab:tableModels)). Soil quality data was provided as
a continuous value between 1 and 5 [Soilgrids dataset, Hengl et al.
-@hengl2017, Poggio et al. -@poggio2021]. The protocol also provided the
elevation, slope and aspect for each of the 200 sites
[@asterscienceteam2019], but no other spatial information such as
coordinates was provided, with the exception of iLand, which used the
blurred coordinates to derive highly detailed soil data.

The simulations were run in the absence of natural disturbances. Only
the model LPJ-GUESS had to include HERE LPJ-GUESS PROVIDE ONE SENTENCE
OF WHY YOU INCLUDED DISTURBANCES AND WHAT TYPE. The simulations were set
up to sample species-specific ingrowth rates per decade and per ha in
the equilibrium state of the model, typically entailing a "spin-up" run
(as we did not provide any forest data). The modeling teams decided on
the simulated area and how they derived these samples. The exact length
of the simulation was also decided by the modeling teams (Table
\@ref(tab:tableModels)). Further details on how each modeling team
prepared the simulations and the outputs are available in SM 2.

@@

The simulations were run for mixed species forests (not multiple
single-species simulations) using mixtures of eleven species or genera
for which ingrowth data of sufficient quality were available from
EuFoRIa: *Fagus sylvatica, Picea abies, Abies alba, Carpinus betulus,
Tilia cordata, Acer pseudoplatanus, Betula spp., Fraxinus excelsior,
Quercus spp., Alnus glutinosa, Pinus sylvestris.* Two models included
their standard set of species for the simulations which is much larger
(i.e., ForClim 1, ForClim 11). This also applies to TreeMig, but in this
model *Betula pendula* was not included because HERE **TREEMIG**
PROVIDES THE END OF THIS SENTENCE SAYING WHY B pendula WAS NOT INCLUDED.
In three models, fewer than these elveven species were simulated (4C,
xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). HERE **4C** PROVIDES
ONE SENTENCE WHY THEY SIMULATED LESS SPECIES. HERE **XCOMP** PROVIDES
ONE SENTENCE WHY THEY SIMULATED LESS SPECIES. HERE **LPJ-GUESS**
PROVIDES ONE SENTENCE WHY THEY SIMULATED LESS SPECIES. Finally, the
model aDGVM2 is based on a suite of broad Plant Functional Types rather
than individual species, and hence this model used its default approach
of PFTs for the simulations.

@@

Three models did not provide results from all the simulated samples or
sites to avoid unrealistic results, as follows. In the model 4C, a
threshold of a maximum basal area of 100 $m^2 ha^{-1}$ was used to avoid
unrealistic stand basal area data, and therefore not all the samples and
sites were reported. The reason for this is that 4C is not suitable for
long-term simulations without management PLEASE **4C** VERIFY THIS IS
WHAT YOU DID. LPJ-GUESS had 2% of the sample outputs as grasslands that
were considered in the data as zero stand basal area and zero ingrowth
for all the tree species simulated. LPJ-GUESS could not simulate one of
the sites HERE **LPJ-GUESS** PROVIDES ONE SENTENCE WHY THEY DID NOT
PROVIDE ALL THE SITES. Landis-II missed 0.2% of the observations HERE
**LANDIS-II** PROVIDES ONE SENTENCE WHY THEY DID NOT PROVIDE ALL THE
SAMPLES/observations.

@@

Both observations and simulated data had no ingrowth in some samples and
at some sites (Table S1). The observed data had slightly more than 3% of
the samples with no ingrowth. Only three models had a larger proportion
of no ingrowth (4C, Landis-II and aDVM2). Three models always simulated
ingrowth for both the 7 and 10 cm threshold (TreeMig and LPJ-GUESS),
i.e., they did not feature any zero values. The other models had a very
low percentage of samples with no ingrowth (0.01% - 1.31%), i.e., they
had distinctly fewer occurrences of zeros compared to the observations.
There was no significant relationship between the mean complexity of the
establishment module (Table \@ref(tab:tableModels)) and the
overestimation proportion at a diameter threshold of 7 cm (SM3).

@@

For both the observed and the simulated data, and both ingrowth
thresholds, species diversity was lower in the ingrowth compared to
stand-level diversity (Figures S3 and S4). Yet, models simulating higher
diversity at the stand level featured higher diversity in ingrowth
(Figure S3 and S4). Most models captured reasonably well or
overpredicted species diversity at the stand level (Figure
\@ref(fig:combinedRichness)A, center; cf. Figure S5), but a
characteristic feature was that simulations had a much lower variability
of diversity compared to observations. Models overpredicting species
richness included ForClim1, ForClim 11, FORMIND and TreeMig (Figure
\@ref(fig:combinedRichness)A, left). Several models underpredicted
species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure
\@ref(fig:combinedRichness)A, right). The majority of the models
captured reasonably well the ingrowth diversity at lower values of
diversity, but most of them strongly underestimated ingrowth diversity
for higher observed diversity levels (Figure
\@ref(fig:combinedRichness)B, cf. Figure S6). Both the models with
feedback and those without feedback from the adult trees to ingrowth
(via seed production) underestimated ingrowth species diversity compared
to the observations. Again, simulated diversity was considerably less
variable than observed diversity (Fig. \@ref(fig:combinedRichness)B).

@@

Second, there was some congruence in the simulation results by species
across the models. The share in the ingrowth by Silver fir (*Abies
alba*) was captured well across the gradient by most models except for
xComp, PICUS, LandClim and Landis-II, which overestimated the ingrowth
share of this species. There were also some peculiarities evident for
some models, such as 4C that did not simulate *Abies alba*. *Fagus
sylvatica* ingrowth was underestimated by many models across the
environmental gradients, including ForCEEPS(f) and FORMIND, which
featured negligible ingrowth at most sites for this species. Most of the
models overestimated the share of *Picea abies* ingrowth in the cold-wet
part of the gradients. *Pinus sylvestris* and *Quercus* spp. represented
a smaller share of the ingrowth in the observed data mostly at warm-dry
sites. Many models, however, erroneously featured ingrowth for these two
species along most of the climatic gradients although at a low basal
area share. Five models (4C, FORMIND, SYBILA, xComp, and LandClim) had
almost no ingrowth of *Pinus sylvestris* and *Quercus* spp., or did not
feature any ingrowth of these species at all.

@@

The species diversity of the entire stand was better captured by the
models than the diversity of
tree ingrowth, and only a few models overpredicted stand-level
diversity. Defining ingrowth composition as being proportional to adult
abundance, regardless of productivity, might be a
simple and conservative assumption to safely maintain relative species
abundances [@hanbury-brown2022], although this may be simplistic.
Furthermore, based on the simulation results there is no evidence that
models with feedback captured better the ingrowth species diversity than
those without feedback. The similar performance of models with and
without feedback is likely because the models are giving more weight to
the regeneration niche arising abiotic and biotic filters than from the
habitat niche of the adult trees [@grubb1977].

@@

The EuFoRIa data as used in this study is unique and highly suited to
better understand tree establishment. However, three aspects of this
data may also represent considerable limitations. First, we made a
comparison of tree ingrowth in a posited equilibrium state, but we
cannot assess how close the empirical data were to such a state. The
data were collected in forest reserves where no management has taken
place for long periods of time. This makes our assumption of an
equilibrium between forest properties and environmental drivers less
haphazard than it might appear at first sight. However, it may explain
why simulated ingrowth rates were higher than the observed ones. Second,
the data were collected for ingrowth above a 7 cm threshold, thus
limiting the assessment of tree establishment in a holistic sense. This
constituted a hard limit based on which we can understand some aspects
of tree establishment, which in its entirety often comprises a rather
long period since germination [@price2001]. However, many environmental
constraints are acting on young trees [@käber2021] that we were unable
to assess. Yet, ingrowth data with lower thresholds are simply not
available in a harmonized manner across large environmental gradients.
Third, the empirical data were collected from rather small plots, while
we sampled simulated ingrowth rates from 1 ha areas, which may lead to
an incorrect representation of space. Even though this strategy it is
not ideal, it represents a common challenge when harmonizing diverse
data sources originating from varied sampling strategies [@portier2022].
It would have been extremely challenging for such a variety of models to
follow a protocol where the spatial sampling size was different at each
of the 200 sites.

The design of our sampling protocol did not include spatial aspects. We
considered a wide range of models of forest dynamics, from stand to
global scale. However, the simulation setup was limited to 1 ha ingrowth
values. The lack of spatial considerations potentially puts landscape
models at a disadvantage, as they have been built to be accurate at the
landscape level. Without spatial context, we are limiting ingrowth to
the seed influx from the stand itself, if the model does not have a
background seed input. However, Global Dynamic Vegetation Models should
not be in disadvantage due to the limited spatial considerations [but
cf. Snell et al. -@snell2014] as they are usually lacking dispersal
between cells as they are based on an abstraction of horizontal space,
or even the trees themselves [@hanburybrown_forest_2022].

@@

Models of forest dynamics are important tools in science and decision
support, and how tree establishment is defined has strong implications
for simulated forest properties. The 15 models and variants used here
are facing similar challenges when representing tree establishment. They
generally overestimate tree establishment rates, and the simulated
establishment niche is not always captured accurately as a function of
the biotic (light) and abiotic (temperature and moisture).

However, most models properly capture the diversity of the initial tree
community, but variations in model formulations, e.g., the presence or
absence of feedback from the adult trees, did not have a strong effect
for capturing the species composition of ingrowth.
